\documentclass[a4paper,10pt,twoside]{article}

%===========PACOTES
\usepackage[body={170mm,235mm}]{geometry}
%\usepackage[portuguese]{babel}
\usepackage{a1}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc} %permite o uso de acentos
%\usepackage[dvips]{color}
\usepackage{amsfonts,amssymb}
%\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{graphicx}	

% makeidx
\usepackage{makeidx}
% make index
\makeindex
%\usepackage[pdftex]{graphicx}


\def\mapright#1#2#3{\smash{\mathop{\hbox to
#3{\rightarrowfill}}\limits^{#1}_{#2}}}

\def\mapleft#1#2#3{\smash{\mathop{\hbox to
#3{\leftarrowfill}}\limits^{#1}_{#2}}}

\def\mapright#1#2{\smash{\mathop{\hbox to 0.90cm{\rightarrowfill}}\limits^{#1}_{#2}}}
\def\mapleft#1#2{\smash{\mathop{\hbox to 0.90cm{\leftarrowfill}}\limits^{#1}_{#2}}}

\def\mapleftright#1#2{\smash{\mathop{\hbox to 0.80cm{\leftarrowfill \rightarrowfill}}\limits^{#1}_{#2}}}
\def\ext{\times \! \vrule depth0pt height5pt width0.35pt}

\def\H{\mathcal H}
\def\D{\mathcal D}
\def\B{\mathcal B}
\def\C{\mathbb C}
\def\R{\mathbb R}
\def\S{\mathbb S}			
\def\U{\mathcal U}
\def\Z{\mathbb Z}

\title{PL-embedding the dual of $J^2$-gems into $\mathbb{S}^ 3$ by an $O(n^ 2)$-algorithm
\footnote{2010 Mathematics Subject Classification: 
57M25 and 57Q15 (primary), 57M27 and 57M15 (secondary)}} 
\author{Sóstenes Lins and Ricardo Machado}

\date{\today}


\begin{document}

\maketitle
%\input{global}
%\input{C.introduction}
%\input{D.sect1.gems}
% \input{sect1}
% \input{sect2}
% \input{sect3}
% \input{sect4}
% \input{sect5}



\begin{abstract}
Let be given a {\em colored 3-pseudo-triangulation} $\mathcal{H}^\star$ with $n$ tetrahedra.
Colored means that each tetrahedron have vertices 
distinctively colored 0,1,2,3.
In a {\em pseudo} 3-triangulation the intersection of simplices might be subsets of simplices
of smaller dimensions, instead of singletons of such faces, 
as for true triangulations. If $\mathcal{H}^\star$ is the dual of a $J^2$-gem (shortly defined),
then we show that $|\mathcal{H}^\star|$ is $\mathbb{S}^3$ and we make available an $O(n^2$)-algorithm to produce a
PL-embedding (\cite{rourke1982introduction}) of $\mathcal{H}^\star$ into $\mathbb{S}^3$.
This is rather surprising because such PL-embeddings are often of exponential size.
This work is the first step towards obtaining, via an $O(n^ 2)$-algorithm,
a framed link presentation inducing the same closed orientable 3-manifold
as the one given by a colored pseudo-triangulation. 
% A {\em $J^ 2$-gem} is a 4-regular, 4-edge-colored planar graph $\mathcal{H}$ obtained from the
% intersection pattern of two Jordan curves $X$ and $Y$ with $2n$ transversal
% crossings.
% These crossings define consecutive segments of $X$ alternatively 
% inside $Y$ and outside $Y$. Color the first type 2 and the second type 3.
% The crossings also define consecutive segments of $Y$ alternatively 
% inside $X$ and outside $X$. Color the first type 0 and the second type 1.
% This defines a 4-regular 4-edge-colored graph $\mathcal{H}$ where the vertices are the crossings
% and the edges are the colored colored segments.
% Let $\mathcal{H}^ \star$ be the 3-dimensional abstract 3-complex formed by taking a set of 
% vertex colored tetrahedra in 1--1 correspondence with $V(\mathcal{H})$ so that each tetrahedra
% has vertices of colors 0,1,2,3. For each $i$-colored edge of $\mathcal{H}$ with ends $u$ and $v$
% paste the corresponding tetrahedra $\nabla_u$ and $\nabla_v$ so as to paste the two triangular
% faces that do not contain a vertex of color $i$ in such a way as to
% to match vertices of the other three colors. 
% We show that the topological space 
% $|\mathcal{H}^\star|$ induced by $\mathcal{H}^\star$ is $\mathbb{S}^ 3$. Moreover we describe an $O(n^2)$-algorithm
% to make available a PL-embedding (\cite{rourke1982introduction}) of $\mathcal{H}^\star$ 
% into $\mathbb{S}^3$. We get explicit coordinates in $\mathbb{S}^ 3$ for the 0-simplices and
% the p-simplices $(p \in \{1,2,3\}$) are linear simplices in the spherical geometry.
Previous work on this topic appear in
\cite{linsmachadoA2012}, \cite{linsmachadoB2012} and \cite{linsmachadoC2012}.
However, the exposition and the new proofs of this paper are meant to be entirely self-contained.

\end{abstract}


\section{Introduction}
\subsection{$J^2$-gems}
A {\em $J^ 2$-gem} is a 4-regular, 4-edge-colored planar graph $\mathcal{H}$ obtained from the
intersection pattern of two Jordan curves $X$ and $Y$ with $2n$ transversal
crossings.
These crossings define consecutive segments of $X$ alternatively 
inside $Y$ and outside $Y$. Color the first type 2 and the second type 3.
The crossings also define consecutive segments of $Y$ alternatively 
inside $X$ and outside $X$. Color the first type 0 and the second type 1.
This defines a 4-regular 4-edge-colored graph $\mathcal{H}$ where the vertices are the crossings
and the edges are the colored colored segments.
Let $\mathcal{H}^\star$ be the 3-dimensional abstract 3-complex formed by taking a set of 
vertex colored tetrahedra in 1--1 correspondence with the set of vertices of $\mathcal{H}$,
$V(\mathcal{H})$, so that each tetrahedra
has vertices of colors 0,1,2,3. This vertex coloring induces a face coloring of the 
triangular faces of the tetrahedron: color $i$ the face opposite to the vertex colored $i$.
For each $i$-colored edge of $\mathcal{H}$ with ends $u$ and $v$
paste the corresponding tetrahedra $\nabla_u$ and $\nabla_v$ so as to paste the two triangular
faces that do not contain a vertex of color $i$ in such a way as to
to match vertices of the other three colors. 
We show that the topological space 
$|K|$ induced by $\mathcal{H}^\star$ is $\mathbb{S}^ 3$. Moreover we describe an $O(n^2)$-algorithm
to make available a PL-embedding (\cite{rourke1982introduction}) of $\mathcal{H}^\star$ 
into $\mathbb{S}^3$. We get explicit coordinates in $\mathbb{S}^ 3$ for the 0-simplices and
the p-simplices $(p \in \{1,2,3\}$) are linear simplices in the spherical geometry.


\subsection{Gems and their duals}
A {\em (3+1)-graph} \index{(3+1)-graph} $\mathcal{H}$ is a connected regular graph of degree 4 where 
to each vertex there are four incident differently colored edges in the color 
set $\{0,1,2,3\}$.
\index{residue}
For $I \subseteq \{0,1,2,3\}$, an {\em $I$-residue} is a component of the subgraph 
induced by the $I$-colored edges. Denote by $v(\mathcal{H})$ the number of
$0$-residues (vertices) of $\mathcal{H}$. For $0\le i < j\le3$, an $\{i,j\}$-residue is also 
called an {\em $ij$-gon} or an $i$- and $j$-colored {\em bigon} (it is an even polygon, where the edges are 
alternatively colored $i$ and $j$). Denote by $b(\mathcal{H})$ the total number of $ij$-gons 
for $0\le i < j\le3$. Denote by $t(\mathcal{H})$ the total number of 
$\overline{i}$-residues for $0\le i\le 3$, where $\overline{i}$ 
means complement of $\{i\}$ in $\{0,1,2,3\}$.

We briefly recall the definition of gems taken from \cite{lins1995gca}.
A {\em $3$-gem} \index{3-gem} is a $(3+1)$-graph $\mathcal{H}$
satisfying $v(\mathcal{H})+t(\mathcal{H})=b(\mathcal{H})$. This relation is equivalent to having the vertices, 
edges and bigons 
restricted to any $\{i,j,k\}$-residue inducing a plane graph where the 
faces are bounded by the 
bigons. Therefore we can embed each such $\{i,j,k\}$-residue into a sphere 
$\mathbb{S}^2$. We consider
the ball bounded this $\mathbb{S}^2$ as induced by the  $\{i,j,k\}$-residue. For this reason
an $\{i,j,k\}$-residue in a 3-gem, $i<j<k$, is also called a \index{triball} {\em triball}.
An $ij$-gon appears once in the boundary of triball $\{i,j,k\}$ and once in the boundary
of triball $\{i,j,h\}$. By pasting the triballs along disks bounded by all 
the pairs of $ij$-gons, $\{i,j\} \subset \{0,1,2,3\}$ of a gem $\mathcal{H}$,
we obtain a closed 3-manifold denoted by $|\mathcal{H}|$. 
This general construction is dual to the one 
exemplified in the abstract and produces any closed 3-manifold. 
The manifold is orientable if and only if 
$\mathcal{H}$ is bipartite, \cite{lins1985graph}. A {\em crystallization} is a
gem which remains connected after deleting all the edges of any given color, that is,
it has one $\{i,j,k\}$-residue for each trio of colors $\{i,j,k\} \subset \{0,1,2,3\}$.

Let $\mathcal{H}^\star$ be the dual of a gem $\mathcal{H}$.
An $\overline{i}$-residue of $\mathcal{H}$ corresponds in $\mathcal{H}^\star$ to a $0$-simplex of $\mathcal{H}^\star$. Most 0-simplices of $\mathcal{H}^\star$
do not correspond to $\overline{i}$-residues of $\mathcal{H}$.
An $ij$-gon of a gem $\mathcal{H}$ corresponds in $\mathcal{H}^\star$ to a {\em PL1-face} formed by a sequence of
1-simplices of $\mathcal{H}^\star$; this PL1-face is the intersection of two PL2-faces of colors $i$ and $j$; their
two bounding $0$-simplices correspond to an $\overline{h}$- and to a $\overline{k}$-residue, where
$\{h,i,j,k\}=\{0,1,2,3\}$.
An $i$-colored edge of $\mathcal{H}$ corresponds to a {\em PL2-face} 
which is a 2-disk triangulated by a subset of $i$-colored 2-simplices of $\mathcal{H}^\star$. Finally to a vertex of $\mathcal{H}$, 
it corresponds a {\em PL3-face} of $\mathcal{H}^\star$ which is a 3-ball formed by a subset of 3-simplices of $\mathcal{H}^\star$.

\begin{proposition}
\label{prop:j2shere}
The $3$-manifold induced by a $J^2$-gem $\mathcal{H}$ is $\mathbb{S}^3$.
\end{proposition}
\begin{proof}
Removing from $\mathcal{H}$ all the edges of a given color still yields a connected graph
which a plane graph and they come embedded so that the faces are the 2-residues.
So $\mathcal{H}$ has four 3-residues, one of each type. 
Denote by $b_{ij}$ the number of $ij$-gons of $\mathcal{H}$.
Each one of these residues are planar graphs
having $v=2n$ vertices, $3v/2$ edges and $b_{12}+b_{13}+b_{23}$, 
$b_{02}+b_{03}+b_{23}$, $b_{01}+b_{13}+b_{03}$ and  $b_{12}+b_{01}+b_{02}$ faces
for, respectively, the $\overline{0}$-, $\overline{1}$-, $\overline{2}$-, 
$\overline{3}$-residue. Adding the four formulas for the Euler characteristic of the sphere
imply that $v(\mathcal{H})+4=b(\mathcal{H})$. Therefore, $\mathcal{H}$ is a crystallization having one $0i$-gon and
one $jk$-gon. This implies that the fundamental group of the induced manifold is trivial:
as proved in \cite{lins1988fundamental}, the fundamental group of the space induced by a crystallization
is generated by $b_{0i}-1$ generators, and in our case this number is 0.
Since Poincaré Conjecture is now proved, we are done. 
However, we can avoid using this fact and, as a bonus, obtaining the validity
of the next corollary, which is used in the sequel.

Assume that $\mathcal{H}$ is a $J^2$-gem which does not induce $\mathbb{S}^3$ and has the smallest possible
number of vertices satisfying these assumptions. By planarity we must have a pair of edges of
$\mathcal{H}$ having the same ends $\{p,q\}$. Consider the graph $\mathcal{H} fus \{p,q\}$ obtained from
$\mathcal{H}$ by removing the vertices $p$, $q$ and the 2 edges linking them as well as welding the 2
pairs of pendant edges along edges of the same color. In \cite{lins1985simple} S. Lins proves that
if $\mathcal{H}$ is a gem, $\mathcal{H}'=\mathcal{H} fus \{p,q\}$ is also a gem and that two exclusive 
relations hold regarding $|\mathcal{H}|$ and $ |\mathcal{H}'|$, their induced 3-manifolds:
either $|\mathcal{H}| = |\mathcal{H}'|$ in the case that $\{p,q\}$ induces a 2-dipole
or else  $|\mathcal{H}| = |\mathcal{H}'| \# (\mathbb{S}^2\times \mathbb{S}^1)$. Since
$\mathcal{H}'$ is a $J^2$-gem, by our minimality hypothesis on $\mathcal{H}$
the valid alternative is the second. But this
is a contradiction: the fundamental group of $|\mathcal{H}|$ would not be trivial, 
because of the summand $\mathbb{S}^2\times \mathbb{S}^1$. 
\end{proof}


\subsection{Dipoles, pillows and balloons} 

Suppose there are $m$ edges linking vertices $x$ and $y$ of a gem, $m \in \{1,2,3\}$.
We say that $\{u,v\}$ is an {\em $m$-dipole} if removing all edges in the colors of the ones linking $x$ to $y$,
these vertices are in distinct components of the graph induced by the edges in the complementary
set of colors. To {\em cancel the dipole} means deleting the subgraph induced by $\{u,v\}$ and identify pairs 
of the hanging edges along the same remaining color. To {\em create the dipole} is the inverse operation.
It is simple to prove that the manifold of a gem is invariant under dipole cancellation or creation.
Even though is not relevant for the present work state the foundational result on gems: two $3$-manifolds
are homeomorphic if and only if any two gems inducing it are linked by a finite number of
cancellations and creations of dipoles, \cite{ferri1982crystallisation, lins2006blobs}.

The dual of a $2$-dipole $\{u,v\}$, with internal colors $i,j$ is named a {\em pillow}. 
It consistes of two PL3-faces $\nabla_u$ and $\nabla_v$ sharing two PL2-faces colored $i$ and $j$.
The {\em thickening of a 2-dipole into a 3-dipole} is defined as follows. Let
$i,j$ be the two colors internal to 2-dipole $\{u,v\}$ and $k$ a third color.
Let $a$ be the $k$-neighboor of $x$ and $b$ be the $k$-neighboor of $y$. Remove
edges $[a,x]$ and $[b,y]$ and put back $k$-edges $[u,v]$ and $[r,s]$. This completes
the thickening. It is simple to prove that the thickening in a gem produces a gem. 
We must be carefull because the inverse blind inverse operation {\em thinning a 3-dipole} 
not always produces a gem. The catch is that the result of the thinning perhaps is not a 2-dipole.
In this sense the thinning move is not local: we must make sure that the result is a 2-dipole.
To the data needed in thinning a 3-dipole $\{u,v\}$ with internal colors $\{i,j,k\}$ 
we must add the $k$-edge $[r,s]$. Note that the $k$-edges $[u,v]$ and $[r,s]$ are in the same $hk$-gon,
where $h$ is the fourth color. Denote by $\Delta_{rs}$ the dual of $[r,s]$.
Let $\nabla_u \cup \nabla v \cup \Delta_{rs}$ be called a {\em balloon}. 
Note that it consists of 2 PL3-faces $\nabla_u$ and $\nabla_v$ sharing three PL2-faces 
in colors $\{i,j,k\}$ together
with a $k$-colored PL2-face whose intersection with  $\nabla_u \cup \nabla_v$ is 
a PL1-face corresponding to the dual of the $hk$-gon, where $h$ is the fourth color.
Let $\nabla_u \cup \nabla_v$ be the {\em balloon's head} and let $\Delta_{rs}$
be the {\em balloon's tail}.


\subsection{The Strategy for obtaining the PL-embedding of the dual of a $J^2$-gem}
We want to find a PL-embedding for the dual, $\mathcal{H}^\star$, of a $J^2$-gem $\mathcal{H}$ 
into $\mathbb{S}^3$. To this end we remove one PL3-face of $\mathcal{H}^\star$ find
in $\mathbb{R}^3$ a PL-embedding of the complementary ball so that their union form
PL-triangulated tetrahedron. After we use the inverse of a stereographic projection with center
in the exterior of the triangulated tetrahedron. In this way we recover in $\mathbb{S}^3$
the missing PL3-face.

In this work we describe the PL-embedded PL3-faces of $\mathcal{H}^\star$ 
into $\mathbb{R}^3$ by making it geometrically clear that its boundary
is a set of 4 PL2-faces, one of each color, forming an embedded $\mathbb{S}^2$ whose interior is
disjoint from the interior of $\mathbb{S}^2$'s corresponding to others PL3-faces. Thus, for our purposes
it will be only necessary to embed the 2-skeleton of $\mathcal{H}^\star$.

A direct approach to find the PL-embedding of the dual of a general $J^2$-gem with $2n$ vertices,
seems very hard.
We split the algorithm into 4 phases. 

In the first phase we find a sequence of $n-1$ 2-dipole thickenings into
3-dipoles so that the final gem is simply a circular arrangement of $n$ 3-dipoles with internal
colors $0,1,2$.
Such a canonical $n$-gem is named a {\em bloboid} and is denoted $\mathcal{B}_n$. 
This indexing decreasing sequence is easily obtainable from the
primal objects, in the case, simplifying $J^2B$-gems: 
$$\mathcal{H}=\mathcal{H}_{n} \mapright{2dip}{thick_1} \mathcal{H}_{n-1} \mapright{2dip}{thick_2} \ldots 
\mapright{2dip}{thick_{n-1}} \mathcal{H}_{1}=\mathcal{B}_n=\mathcal{B}.$$ 

In the second phase we first find (phase 2A) specific abstract PL-triangulations, 
for the PL2-faces for the index increasing sequence of abstract colored 2-dimensional PL-complexes.
Each of these complexes, later, are going to be PL-embedded into $\mathbb{R}^3$ so 
that the PL2-faces are topologically 2-spheres with disjoint interior. Attaching 
3-balls bounded by these spheres we get the dual of the $J^2$-gem $\mathcal{H}$ (with a vertex
removed).
$$\mathcal{B}^\star=\mathcal{H}^\star_{1} \mapright{bp}{move_1}  \mathcal{H}^\star_{2}  
\mapright{bp}{move_{2}} \ldots
\mapright{bp}{move_{n-1}} \mathcal{H}^\star_{n}=\mathcal{H}^\star.$$ 
In parallel to the construction of the sequence $\mathcal{H}^\star_{m}$'s we also construct
(phase 2B) a sequence of {\em wings} $\mathcal{W}^\star_{m}$'s and their {\em nervures} 
$\mathcal{N}^\star_{m}$'s so that  $\mathcal{W}^\star_{m} \cup \mathcal{N}^\star_{m}$ 
is an easy to define adequate planar graph:
$$\mathcal{W}^\star_{1} \cup \mathcal{N}^\star_{1} \mapright{wbp}{move_{1}} 
\mathcal{W}^\star_{2} \cup \mathcal{N}^\star_{2} \mapright{wbp}{move_{2}} \ldots
\mapright{wbp}{move_{n-1}} 
\mathcal{W}^\star_{n} \cup \mathcal{N}^\star_{n}=
\mathcal{W}^\star \cup \mathcal{N}^\star.$$ 
Each wing $\mathcal{W}_m^\star$'s corresponds to a section 
of the previous sequence $\mathcal{H}_m^\star$'s by two adequated 
fixed semi-planes $\Pi_\ell$ and $\Pi_r$.
The construction of the planar graphs $\mathcal{W}_m^\star \cup \mathcal{N}_m^\star$ is
recursive. Initially $\mathcal{W}_1^\star$ is a pincel of lines and $\mathcal{N}_1^\star$ 
is $\varnothing$.  Going from $\mathcal{W}_{m-1}^\star \cup \mathcal{N}_{m-1}^\star$ 
to $\mathcal{W}_{m}^\star \cup \mathcal{N}_{m}^\star$ is very simple: two new vertices and four
new edges appear, so as to maintain planarity.
% a vertex is broken into two
% and a new vertex of $\mathcal{N}_{m}^\star$ appear;
% two new edges are adjoined to $\mathcal{W}_{m}^\star$ and two new edges are adjoined to 
% $\mathcal{N}_{m}^\star$, in a way as to respect some cyclic ordering around the new vertices.


In the third phase we make the abstract final element 
$\mathcal{W}^\star \cup \mathcal{N}^\star$
of the second phase {\em rectilinearly} (that is each edge is a straihgt line segment) PL-embedded. 
By a cone construction we obtain from the rectilinearly PL-embedded 
$\mathcal{W}^\star \cup \mathcal{N}^\star$  a special
PL-complex, named $\mathcal{H}_1^\diamond$. This complex does not correpond to a gem dual
and it can be loosely explained as $\mathcal{H}_1^\star$ with all balloon's heads ``opened''.
% More details can only be given in the actual construct of Section \ref{}.


The fourth phase, the {\em pillow filling phase} 
starts with $\mathcal{H}_1^\diamond$ and the uses the abstract sequence 
$\mathcal{H}^\star_{m}$'s to produce 
a pillow filling sequence $$\mathcal{H}^\diamond_{1}  \mapright{pillow}{filling_{1}}  
\mathcal{H}^\diamond_2  \mapright{pillow}{filling_{2}} \ldots  \mapright{pillow}{filling_{n-1}}
\mathcal{H}^\diamond_{n}=\mathcal{H}^\star.$$
In this phase everything is embedded into $\mathbb{R}^3$ and the last element
is the PL-embedding that we seek. 
% It is worth mentioning that (except for the last one) these
% complexes have no counterpart in the dual gems, that is they do not have PL2-faces, only
% the part correspondig to the balloon's head  $\nabla_u \cup \nabla_v$.

The whole procedure can be implemented as a formal 
algorithm that takes $O(n^2)$-space
and $O(n^2)$-time complexity, where $2n$ is the number of vertices of the original $J^2$-gem.

\subsection{Acknowledgements}
We thank the anonimous referees, which made us think and provide an $O(n)$-solution
for finding the wing sequence, instead of an $O(n^3)$-solution based on a bare Tutte's
barycetric methods.

\section{Details of the whole construction}
\subsection{First phase: from a $J^2$-gem $\mathcal{H}$ to a bloboid $\mathcal{B}$}

\begin{proposition}
\label{prop:startingwith}
Starting with a $J^2$-gem $\mathcal{J}^2$ with $2n$ vertices we can arrive to an
$n$-bloboid $\mathcal{B}$ by means of $n-1$ operations which thickens a 2-dipole into a 3-dipole,
producing a sequence of $J^2B$-gems each inducing $\mathbb{S}^3$,
$$(\mathcal{J}^2=\mathcal{H}_{n},\mathcal{H}_{n-1},\ldots, \mathcal{H}_{2},\mathcal{H}_{1}=\mathcal{B}).$$

\end{proposition}
\begin{proof}
The proof is by induction. For $\ell=n$ we have $\mathcal{H}_{n}=\mathcal{J}^2$ and
so it is a $J^2B$-gem, establishing the basis of the induction. Assume that
$\mathcal{H}_\ell$ is a $J^2B$-gem.
For $\ell>1$, let $\mathcal{H}'_\ell$ denote  $\mathcal{H}_\ell$ after cancelling the blobs.
Since $\mathcal{H}'_\ell$ is a $J^2$-gem by the Jordan curve theorem
a 2-dipole is present in it. The same 2-dipole is also present in $\mathcal{H}_\ell$.
Therefore it can be thickened, defining  $J^2B$-gem $\mathcal{H}_{\ell-1}$, which
establishes the inductive step. In face of Proposition \ref{prop:j2shere} and from
the fact that thickening dipoles on gems produce gems inducing the same manifold, 
every member of the sequence induces $\mathbb{S}^3$.
\end{proof}

% The above proof is illustrated in Fig. \ref{fig:j2gemrSEQ}. Note that the thickenings are of two
% types: either of type $0$ where the edges $e$ and $f$ are $0$-edges in the interior of the $23$-gon, or else
% of type $1$ if they  are $1$-edges in its exterior. The above sequence is by no means unique because
% various 2-dipole choices present themselves along the way. We fix one and such a sequence and 
% used it as the basis for
% establishing bounds on the size of the colored complexes, 
% dual of the $J^2$-gem. This is done in the companion
% paper \cite{linsmachadoB2012}.

\subsection{Second phase: colored abstract complexes, their wings and nervures}
\subsection{Third phase: a rectilinear PL-embedding for 
$\mathcal{W}^\star \cup \mathcal{N}^\star$ and its induced $\mathcal{H}_1^\diamond$}
\subsection{Fourth phase: filling the pillows to obtain the PL-embedding of $\mathcal{H}^\star$ we seek}

\section{Conclusion and further work}

\include{E.refAffiliation}
%\input{affiliation}


\end{document}



